Using Occupancy Theory and Hierarchical Models to Estimate Otter Population in Southeastern Minnesota

### 2010-02-28 Paper.tex
Edited appendix and table

    \subsection{Testing the independence assumption}
    \label{shat}
    Since one of the assumptions of occupancy modeling is that there is 
	independence between sites, we wanted to see if this assumption held true 
	for our data.  In order to do this, we calculated the statistic \(S\), which
	we define as the number of successive \(0\), \(1\)'s minus the number of
    successive \(1\), \(1\)'s per flight. Under the null hypothesis that each
    site is an independent Bernoulli trial with probability of a \(1 = p\) we were able
    to compute the mean and standard deviation of \(S\), which we used to
    calculate a Z-score, where \(Z = (S – E[S])/SD[S]\). In order to compute these values, we needed to find the
    expected value and variance of each term in \(S\). These caclulations can be
	found in the appendix.

@@

    % Mississippi River!
    \begin{center}
    \begin{tabular}{|l|l|l|l|l|l|}
        \hline
        \multicolumn{6}{|l|}{\textbf{Mississippi River}} \\
        \hline
        \multirow{2}{*}{Year} & \multirow{2}{*}{Size} & \multirow{2}{*}{Flights}
        & \multicolumn{2}{|l|}{\# Sig. Z-Scores} &
        \multirow{2}{*}{\(E[|z_{all}|-|z_{alt}|]\)} \\
        \cline{4-5}
        & & & All Plots & Alt. Plots & \\
        \hline
        \multirow{3}{*}{2003} & 402m & \multirow{3}{*}{20} & 7 & 0 & 0.915 \\
        \cline{2-2} \cline{4-6}
        & 804m & & 5 & 0 & 0.651 \\
        \cline{2-2} \cline{4-6}
        & 1608m & & 1 & 0 & 0.383 \\
        \hline
        \multirow{3}{*}{2004} & 402m & \multirow{3}{*}{6} & 5 & 3 & 0.625 \\
        \cline{2-2} \cline{4-6}
        & 804m & & 3 & 0 & 1.428 \\
        \cline{2-2} \cline{4-6}
        & 1608m & & 2 & 0 & 0.053 \\
        \hline
    \end{tabular}
    \end{center}

    % Whitewater River!
    \begin{center}
    \begin{tabular}{|l|l|l|l|l|l|}
        \hline
        \multicolumn{6}{|l|}{\textbf{Whitewater River}} \\
        \hline
        \multirow{2}{*}{Year} & \multirow{2}{*}{Size} & \multirow{2}{*}{Flights}
        & \multicolumn{2}{|l|}{\# Sig. Z-Scores} &
        \multirow{2}{*}{\(E[|z_{all}|-|z_{alt}|]\)} \\
        \cline{4-5}
        & & & All Plots & Alt. Plots & \\
        \hline
        \multirow{2}{*}{2003} & 402m & \multirow{2}{*}{15} & 6 & 2 & 0.722 \\
        \cline{2-2} \cline{4-6}
        & 804m & & 4 & 1 & 0.634 \\
        \hline
        \multirow{2}{*}{2004} & 402m & \multirow{2}{*}{6} & 3 & 1 & 1.189 \\
        \cline{2-2} \cline{4-6}
        & 804m & & 2 & 2 & 0.612 \\
        \hline
    \end{tabular}
    \end{center}

    % Zumbro River!
    \begin{center}
    \begin{tabular}{|l|l|l|l|l|l|}
        \hline
        \multicolumn{6}{|l|}{\textbf{Zumbro River}} \\
        \hline
        \multirow{2}{*}{Year} & \multirow{2}{*}{Size} & \multirow{2}{*}{Flights}
        & \multicolumn{2}{|l|}{\# Sig. Z-Scores} &
        \multirow{2}{*}{\(E[|z_{all}|-|z_{alt}|]\)} \\
        \cline{4-5}
        & & & All Plots & Alt. Plots & \\
        \hline
        \multirow{2}{*}{2003} & 402m & \multirow{2}{*}{11} & 10 & 6 & 1.883 \\
        \cline{2-2} \cline{4-6}
        & 804m & & 8 & 5 & 1.287 \\
        \hline
        \multirow{2}{*}{2004} & 402m & \multirow{2}{*}{4} & 3 & 3 & 2.425 \\
        \cline{2-2} \cline{4-6}
        & 804m & & 3 & 1 & 2.343 \\
        \hline
    \end{tabular}
    \end{center}

@@

%Needs to be fixed/formatted better... I didn't know how to make an appendix
\appendix
	We let \(A = \text{\# of
    0-1's}\), \(B=\text{\# of 1-1's}\) and \(p=P(``1")\text{ in the sequence}\).
    In order to find \(E[s]\), we first needed to find \(E[A]\) and
    \(E[B]\).

>>>>

    \subsection{Testing the independence assumption}
    \label{shat}
    Since one of the assumptions of occupancy modeling is that there is 
	independence between sites, we wanted to see if this assumption held true 
	for our data.  In order to do this, we calculated the statistic \(S\), which
	we define as the number of successive \(0\), \(1\)'s minus the number of
    successive \(1\), \(1\)'s per flight. Under the null hypothesis that each
    site is an independent Bernoulli trial with probability of a \(1 = p\) we were able
    to compute the mean and standard deviation of \(S\), which we used to
    calculate a Z-score, where \(Z = (\hat{S} - E[S])/SD[S]\). In order to compute these values, we needed to find the
    expected value and variance of each term in \(S\). These caclulations can be
	found in the appendix \ref{math}.

@@

    % Mississippi River!
    \begin{center}
    \begin{tabular}{|l|l|l|l|l|}
        \hline
        \multicolumn{5}{|l|}{\textbf{Mississippi River}} \\
        \hline
        \multirow{2}{*}{Year} & \multirow{2}{*}{Size} & \multirow{2}{*}{Flights}
        & \multicolumn{2}{|l|}{\# Sig. Z-Scores} \\
        \cline{4-5}
        & & & All Plots & Alt. Plots \\
        \hline
        \multirow{3}{*}{2003} & 402m & \multirow{3}{*}{20} & 7 & 0 \\
        \cline{2-2} \cline{4-5}
        & 804m & & 5 & 0 \\
        \cline{2-2} \cline{4-5}
        & 1608m & & 1 & 0 \\
        \hline
        \multirow{3}{*}{2004} & 402m & \multirow{3}{*}{6} & 5 & 3 \\
        \cline{2-2} \cline{4-5}
        & 804m & & 3 & 0 \\
        \cline{2-2} \cline{4-5}
        & 1608m & & 2 & 0 \\
        \hline
    \end{tabular}
    \end{center}

    % Whitewater River!
    \begin{center}
    \begin{tabular}{|l|l|l|l|l|}
        \hline
        \multicolumn{5}{|l|}{\textbf{Whitewater River}} \\
        \hline
        \multirow{2}{*}{Year} & \multirow{2}{*}{Size} & \multirow{2}{*}{Flights}
        & \multicolumn{2}{|l|}{\# Sig. Z-Scores} \\
        \cline{4-5}
        & & & All Plots & Alt. Plots \\
        \hline
        \multirow{2}{*}{2003} & 402m & \multirow{2}{*}{15} & 6 & 2 \\
        \cline{2-2} \cline{4-5}
        & 804m & & 4 & 1 \\
        \hline
        \multirow{2}{*}{2004} & 402m & \multirow{2}{*}{6} & 3 & 1 \\
        \cline{2-2} \cline{4-5}
        & 804m & & 2 & 2 \\
        \hline
    \end{tabular}
    \end{center}

    % Zumbro River!
    \begin{center}
    \begin{tabular}{|l|l|l|l|l|}
        \hline
        \multicolumn{5}{|l|}{\textbf{Zumbro River}} \\
        \hline
        \multirow{2}{*}{Year} & \multirow{2}{*}{Size} & \multirow{2}{*}{Flights}
        & \multicolumn{2}{|l|}{\# Sig. Z-Scores} \\
        \cline{4-5}
        & & & All Plots & Alt. Plots \\
        \hline
        \multirow{2}{*}{2003} & 402m & \multirow{2}{*}{11} & 10 & 6 \\
        \cline{2-2} \cline{4-5}
        & 804m & & 8 & 5 \\
        \hline
        \multirow{2}{*}{2004} & 402m & \multirow{2}{*}{4} & 3 & 3 \\
        \cline{2-2} \cline{4-5}
        & 804m & & 3 & 1 \\
        \hline
    \end{tabular}
    \end{center}

@@

%Needs to be fixed/formatted better... I didn't know how to make an appendix
\appendix
	\section{Deriving \(E[S]\) and \(SD[S]\)}
	\label{math}
	We let \(A = \text{\# of
    0-1's}\), \(B=\text{\# of 1-1's}\) and \(p=P(``1")\text{ in the sequence}\).
    In order to find \(E[s]\), we first needed to find \(E[A]\) and
    \(E[B]\).
